Bayesian exponential random graph modeling of KWL Burial network

### 2020-06-22 analysis/paper/paper.Rmd
edit method section and add references

We analyzed archaeological burials from a Iron Age site, Kiwulan, in northeastern Taiwan based on data collected from published excavation report, and original fieldwork notes to generate networks. The data includes information of gender, age, burial rites, and types of burial goods. We do not include the burials in Chinese phase in our network analysis because of the small sample size. A network is a structure consists of a set of nodes (actors) and edges (ties) connecting them representing relationships [@Brughmans2018]; in our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common. This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp. 252]. Our purpose is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships. In other words, we explore which variables may have contributed to the observed economic inequality across burials using network analysis approaches, and any changes in network structure before and after the foreign contacts.

We hypothesized that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society. The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006]. In addition to valuable foreign goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals to built their personal status and power. This led to the monopolization of valuable goods among a small numbers of individuals and the society gradually switched to a structure with social differentiation. If the social inequality gradually increased in Kiwulan, then we will observe it from the burials by analyzing the networks generated by the social processes. This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.

## Model specification and estimation
<!--Network construction-->

We use Bayesian inference on exponential random graph models (ERGMs) to quantify the relationships among burials and test our hypothesis. ERGMs are statistical models for building complex social networks where possible ties are assumed as random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. A number of parameters used for modeling the dependencies expresses the structure or property of a network, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with similar attribute), and popularity (actors have many relationships with others) [@Robins2007; @Snijders2006]. With a research question for a certain social process, we can build a hypothesis-based model to generate a distribution of random networks according to the parameters we specified, also called configurations or network statistics. This helps to understand the underlying processes of an observed network by comparing it with a large set of possible networks generated by the model.

Table \@ref(tab:ergm-terms-table) lists the parameters we used in the model construction for burials with archaeological interpretations. Every term in an ERGM has an associated algorithm for computing its value for network [@Morris2008]. Based on our hypothesis, the burial network after the European presence will present less reciprocity, less transitivity, but higher popularity structure effects compared to the network before the European presence. Since burials tend to have multiple types of burial goods in common, such as personal ornaments, imported porcelains, and gold foils, the network ties are weighted instead of a binary data expressed by 0 and 1 [@Snijders2006]. We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods. In addition to wealth, we also consider another two social variables, age and gender of the interred individual, serving as indicators for distinguishing ascribed or achieved status in a society [citation].

@@

We conducted a network analysis using the R programming language with the ergm package for simulation and model specification [@Hunter2008; @Morris2008]. We used the bergm package for Bayesian modeling [@Caimo2014]. <!-- why did we use two packages? why not just one? this needs to be expanded a bit to make it clear to the reader --> We set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.

@@

<!-- "We test our hypothesis by measuring reciprocality,  transitivity, and  popularity structure effects and expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects" This is where you can connect the anthropological question, the archaeological hypothesis and the statistical method for testing the hypothesis  -->

>>>>

We analyzed archaeological burials from a Iron Age site, Kiwulan, in northeastern Taiwan based on data collected from published excavation report, and original fieldwork notes to generate networks. The data includes information of gender, age, burial rites, and types of burial goods. We do not include the burials in Chinese phase in our network analysis because of the small sample size. A network is a structure can be represented as a graph consisting of a set of nodes (actors) and edges (ties) connecting them [@Brughmans2018; @Robins2007]; in our case, burials are nodes linked by edges if any pair of burial has the prestige goods in common. This is based on the assumption that the prestige goods imply social practices of trade, exchange, and gifting that reflect social relations between individuals [@Coward2013, pp. 252]. Since burials tend to have multiple prestige goods in common, such as personal ornaments and imported porcelains, the network ties are weighted instead of binary variables, where the value 1 represents a tie between two nodes and the value 0 otherwise [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors.

Our purpose is to understand the underlying mechanisms that shape the relationships between people at Kiwulan, and the effect of foreign contact on these relationships. In other words, we explore which variables may have contributed to the observed economic inequality across burials using network analysis approaches, and any changes in network structure before and after the foreign contacts. We hypothesized that after the start of the European presence in the 17th century, a group of ambitious individuals competed with each other for access to trade with the Europeans due to the high value of trade goods in their society. The high value of trade goods had been identified by the historical documents written by the Europeans and the Han Chinese [@Borao2009; @LiandWu2006]. In addition to valuable foreign goods, the indirect colonialism in northeastern Taiwan may suggest the potentially manipulation of the European colonial image by local ambitious individuals to built their personal status and power. This led to the monopolization of valuable goods among a small numbers of individuals and the society gradually switched to a structure with social differentiation. If the social inequality gradually increased in Kiwulan, then we will observe it from the burials by analyzing the networks generated by the social processes. This is important to understand how an Indigenous society reacted to the presence of foreign imperial power.

## Model specification 
<!--Network construction-->

We use Bayesian inference on exponential random graph models (ERGMs) to quantify the relationships among burials and test our hypothesis. ERGMs are statistical models for building complex social networks where possible ties are assumed as random variables and dependent on actor variables or the presence or absence of other ties [@Robins2007]. In ERGMs, ties form a small structure in a network called graph configuration that describes the form of dependence, such as reciprocity (relationship between two actors), transitivity (relationship between two actors through a shared third actor), homophily (relationship between actors with similar attribute), or popularity (actors have many relationships with others) [@Robins2007; @Snijders2006; @Morris2008]. Those configurations are the structure or property of a network and can be expressed as network statistics. By specifying the forms of configurations, we can build a hypothesis-based model to generate a distribution of random networks. This distribution consists of a large number of possible networks that provides a statistical inference to an observed network [@Robins2007]. This helps to understand whether our observed network shows significantly more or less property of interest than the large set of random networks we modeled.

Table \@ref(tab:ergm-terms-table) lists the parameters we used in the model construction for burials with archaeological interpretations. Every parameter in an ERGM has an associated algorithm for computing its value for a network [@Morris2008]. We test our hypothesis by measuring reciprocity, transitivity, popularity, and homophily structure effects. We expect that if our hypothesis is a good match with the data then we should see less reciprocally, less transitivity, but higher popularity structure effects after the European presence. We assigned the burials into three classes, high, medium, and low, as an index of wealth, based on the quantity of burial goods. In addition to wealth, we also consider another two social variables, age and gender of the interred individual, serving as indicators for distinguishing ascribed or achieved status in a society [citation].

@@

We analyzed social networks using the R programming language with bergm package for Bayesian exponential random graph models [@Caimo2014]. Bergm uses ergm packages for simulation and model specification, and provides a framework for Bayesian parameter estimation on the result of exponential random graph models [@Hunter2008; @Morris2008; @Handcock2008]. We set the number of chains to six. For each chain, the number of burn-in iterations is 200 and the number of iterations after the burn-in is 2000. We set the number of iterations used to simulate a network y′ at each iteration to 10,000. The models are assessed by goodness-of-fit (GOF) diagnostics in the Bayesian framework, where the observed network is compared with the set of networks simulated from the estimated posterior distribution of the parameters of the model [@Caimo2011; @Caimo2017]. This provides a statistical way to check how well the estimated posterior parameter distribution can reproduce networks similar to the general structural features of the observed network.

@@
